---
title: "Is it possible to make a million on Fiverr?"
subtitle: "Insights from analysing 9 Niches on Fiverr"
output: 
  html_document: default
  md_document: default
params:
  # filter: '{"keyword": {"$in": ["airbnb", "alibaba", "amazon", "amazon fba", "data analysis", "data science", "fiverr", "java programming", "market analysis", "market research", "passive income", "python programming", "r programming", "reporting" , "scraping", "six figures"]}}'
  filter: '{"keyword": {"$in": ["Voiceover", "Copywriting", "Resume Writing", "Press release Writing", "Video Marketing", "Website Building", "Mobile App Development", "Graphic Design", "seo"]}}'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F, fig.height = 10, fig.width = 20)
```




```{r}
MONGO_URI <- "mongodb://localhost/fiverr"
DATABASE <- "fiverr"
cpation.text <- "by Jens Laufer (https://www.fiverr.com/jens_laufer)"
label.color <- "black"
arrow.color <- "#BAB0AC"
label.size <- 7
DATA_FILE <- '../_data/gigs.csv'
```

```{r}
devtools::install_github('bbc/bbplot')
```


```{r}
source('include/data.R',encoding = 'utf-8')
source('include/plots.R', encoding='utf-8')
```



```{r}
example.top.seller.gigs.df <- read.csv('../_data/top_seller_gigs.csv') %>%
  mutate(reviewed_revenue=start_price*reviews_count) 
reported.revenue.df <- read_csv('../_data/reported_revenue.csv')

example.top.seller.revenue.df <- example.top.seller.gigs.df %>% select(real_name, username, gig, reviewed_revenue) %>% 
  group_by(real_name, username) %>%
  summarise(total_reviewed_revenue=sum(reviewed_revenue)) %>% inner_join(reported.revenue.df)
```





```{r}
levels <- (example.top.seller.revenue.df %>% arrange(reported_revenue))$real_name

example.top.seller.revenue.df %>%
  rename(`Total Reviewed Revenue`=total_reviewed_revenue) %>%
  rename(`Reported Revenue`=reported_revenue) %>%
  gather(-username, -real_name, key='revenue_type', value='val') %>%
  ggplot(aes(x=real_name, y=val, group=revenue_type,fill=revenue_type))+
  geom_bar(stat='identity', position = 'dodge')  +
  scale_fill_tableau()+
  scale_x_discrete(limits=levels)+
  coord_flip()+
  bbc_style()
```
```{r}

top_gigs <- example.top.seller.gigs.df %>%
  group_by(username) %>%
  top_n(4, reviewed_revenue) %>%
  ungroup()

levels <- (top_gigs %>% arrange(real_name, reviewed_revenue))$gig

top_gigs %>%
  ggplot(aes(x=gig, y=reviewed_revenue, fill=real_name))+
  geom_bar(stat='identity') +
  scale_fill_tableau() +
  scale_y_continuous(labels=comma, breaks=seq(0, 500000, 50000))+
  scale_x_discrete(limits=levels)+
  coord_flip() +
  bbc_style()
```



```{r}
get.products(DATABASE, MONGO_URI, '{}') %>% as_tibble() %>% 
  mutate(price_start=replace_na(price_start,0), score_num=replace_na(score_num,0), revenue=price_start*score_num,title_length=nchar(title)) %>%
  distinct(title, .keep_all = T) %>%
  head(10) %>%
  ggplot(aes(x=reorder(title, revenue), y=revenue))+
  geom_bar(stat='identity', fill='#4E79A7') +
  scale_y_continuous(labels = comma)+
  coord_flip()+
  bbc_style()
```









The story of [How a dad made almost 1 million on Fiverr](https://www.cnbc.com/2018/04/24/how-this-dad-made-almost-1-million-on-fiverr.html?__source=msn%7Cmoney%7Cinline%7Cstory%7C&par=msn&doc=105300315) inspired me to analyse data from fiverr to gain insights how realistic a story like this. So I was looking for more stories from people to get an idea what they earn and what they are doing. 

| Seller | Type of service | Income |
| ------ | --------------- | -----: |
| [Joel Young](https://www.cnbc.com/2018/04/24/how-this-dad-made-almost-1-million-on-fiverr.html?__source=msn%7Cmoney%7Cinline%7Cstory%7C&par=msn&doc=105300315)| Voice Over | $1,000,0000|
|[Charmaine Pocek](https://www.forbes.com/sites/laurashin/2016/05/31/how-these-3-people-make-6-figures-a-year-on-fiverr/)| Resume writing | $480,000|
| [Levi Newman](https://www.forbes.com/sites/laurashin/2016/05/31/how-these-3-people-make-6-figures-a-year-on-fiverr/)|Copywriting| $180,000 |
| [Redd Horrocks](https://www.forbes.com/sites/laurashin/2016/05/31/how-these-3-people-make-6-figures-a-year-on-fiverr/)|Voice Over| $180,000 |
|[Alex Fasulo](https://www.msn.com/en-us/money/bills-to-pay/how-this-25-year-old-freelancer-made-dollar150000-in-6-months-off-fiverr/ar-AAzkNOA)| Press Release Writing | $150,000 |
|[David Feldman](https://www.cnbc.com/2018/12/18/how-this-38-year-old-is-making-6-figures-freelancing-on-fiverr.html)|Copywriting|$100,000|

I identified more niches with [The 8 highest-paying side hustles on Fiverr](https://www.cnbc.com/2018/04/30/the-8-highest-paying-side-hustles-on-fiverr.html) 

I entered the numbers game for these niches:

- Voiceover
- Copywriting
- Resume Writing
- Press release Writing
- Video Marketing
- Website Building
- Mobile App Development
- Graphic Design
- SEO

